---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pristineR

<!-- badges: start -->
<!-- badges: end -->

`pristineR` is an R package designed to simplify common
data-cleaning and summarization tasks through safer,
user-friendly helper functions.

The package includes tools for parsing inconsistent date formats, generating summary statistics, and cleaning messy column names.

## Why pristineR?

Many real-world datasets contain inconsistent date formats, messy column names, and missing values that require repetitive cleaning before analysis can begin.

`pristineR` was created to simplify these common preprocessing tasks through lightweight helper functions that are easy to use, well documented, and designed to fail safely when given invalid inputs.

## Installation

You can install the development version of pristineR from
[GitHub](https://github.com/) with:

```{r eval = FALSE}
#install.packages("remotes")
#install.packages("devtools")
devtools::install_github("ADC-405-S26/pristineR")
```

## Load the Package

```{r}
library(pristineR)
```

## Example Dataset

The package contains an example dataset called
`messy_survey`.

```{r}
messy_survey
```

The dataset intentionally contains inconsistent column names, mixed date formats, and missing values.

## Clean Column Names

Datasets imported from spreadsheets or surveys often
contain spaces, punctuation, or inconsistent capitalization
in column names.

The `clean_names()` function converts messy
column names into snake_case format.

```{r}
names(messy_survey)

cleaned_data <- clean_names(messy_survey)

names(cleaned_data)
```

## Parse Dates

Datasets frequently contain dates stored in multiple formats.
The `safe_date_parse()` function safely converts these values into standard Date objects.

```{r}
cleaned_data$submission_date <-
  safe_date_parse(cleaned_data$submission_date)

cleaned_data
```

## Handling Invalid Dates

If a date cannot be parsed, `safe_date_parse()` returns
`NA` and produces a warning message instead of stopping
execution completely.

```{r}
safe_date_parse(c("2024-01-01", "bad-date"))
```

## Generate Summary Statistics

The `quick_summary()` function generates common summary
statistics for selected numeric columns.

```{r}
quick_summary(
  cleaned_data,
  c("exam_score")
)
```

The summary includes mean, standard deviation, minimum, maximum, and missing value counts.


## Function Parameters

- `safe_date_parse(x)`
  - `x`: a character vector containing date values

- `quick_summary(data, columns)`
  - `data`: a data frame
  - `columns`: character vector of numeric columns
    to summarize

- `clean_names(data)`
  - `data`: a data frame with column names to clean

## Additional Example

`pristineR` can also be used with other messy datasets.

```{r}
example_data <- data.frame(
  "Customer Name" = c("Alice", "Bob"),
  "Purchase.Date" = c("2024-01-01", "01/15/2024"),
  "Total Sales $" = c(120, 95)
)

example_data <- clean_names(example_data)

example_data$purchase_date <-
  safe_date_parse(example_data$purchase_date)

example_data
```

## Package Functions

`pristineR` currently includes:

- `safe_date_parse()`
- `quick_summary()`
- `clean_names()`
