---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pristineR

<!-- badges: start -->
<!-- badges: end -->

`pristineR` is an R package designed to simplify common
data-cleaning and summarization tasks through safer,
user-friendly helper functions.

The package includes tools for parsing inconsistent date formats, generating summary statistics, and cleaning messy column names

## Installation

You can install the development version of `pristineR`
from GitHub with:

```{r}
# install.packages("devtools")
devtools::install_github("ADC-405-S26/pristineR")
```

## Load the Package

```{r}
library(pristineR)
```

## Example Dataset

The package contains an example dataset called
`messy_survey`.

```{r}
messy_survey
```

## Clean Column Names

The `clean_names()` function converts messy
column names into snake_case format.

```{r}
cleaned_data <- clean_names(messy_survey)

names(cleaned_data)
```

## Parse Dates

The `safe_date_parse()` function safely parses mixed-format
dates into standard Date objects.

```{r}
cleaned_data$submission_date <-
  safe_date_parse(cleaned_data$submission_date)

cleaned_data
```

## Generate Summary Statistics

The `quick_summary()` function generates common summary
statistics for selected numeric columns.

```{r}
quick_summary(
  cleaned_data,
  c("exam_score")
)
```

## Package Functions

`pristineR` currently includes:

- `safe_date_parse()`
- `quick_summary()`
- `clean_names()`
