---
title: "Cleaning Messy Data with pristineR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cleaning Messy Data with pristineR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Data cleaning is one of the most common and repetitive tasks
in data analysis workflows. Real-world datasets often contain
inconsistent date formats, messy column names, and missing values that require a lot of the same tedious cleaning before analysis can begin.

The `pristineR` package provides functions that simplify
common data-cleaning tasks through safer and more user-friendly
workflows.

This vignette demonstrates how to use: `clean_names()`, `safe_date_parse()`, and `quick_summary()` using the included `messy_survey` dataset.

# Load the Package

```{r setup}
library(pristineR)
```

# Example Dataset

The package includes an example dataset called
`messy_survey`.

```{r}
messy_survey
```

The dataset intentionally contains inconsistent column names, mixed date formats, and missing values.

# Cleaning Column Names

Messy column names can make analysis more difficult.
The `clean_names()` function converts names
into standardized snake_case formatting.

```{r}
cleaned_data <- clean_names(messy_survey)

names(cleaned_data)
```

# Parsing Dates

Datasets often contain dates stored in multiple formats.
The `safe_date_parse()` function safely converts these
dates into standard Date objects.

```{r}
cleaned_data$submission_date <-
  safe_date_parse(cleaned_data$submission_date)

cleaned_data
```

# Generating Summary Statistics

The `quick_summary()` function generates common
summary statistics for numeric variables.

```{r}
quick_summary(
  cleaned_data,
  c("exam_score")
)
```

The summary includes: mean, standard deviation, minimum, maximum, and missing 
value counts.

# Complete Workflow - Putting it all Together

The following example demonstrates a complete
cleaning workflow using `pristineR`.

```{r}
cleaned_data <- messy_survey |>
  clean_names()

cleaned_data$submission_date <-
  safe_date_parse(cleaned_data$submission_date)

summary_table <- quick_summary(
  cleaned_data,
  c("exam_score")
)

summary_table
```

# Conclusion

The `pristineR` package simplifies several common
data-cleaning tasks that frequently arise in R workflows.

Raw datasets are often difficult to analyze immediately due
to inconsistent formatting and missing values. By combining
the functions in `pristineR`, users can quickly transform
messy data into a cleaner and more analysis-ready format.
